---
title: "R Notebook"
output: html_notebook
---

Day 1 task 1: count the number of times a depth measurement increases from the previous measurement.

First, let's read in the data:

```{r}
depth_meas <- as.numeric(readLines("data/input1.txt"))
```

Calculate difference between subsequent rows and sum the positive ones:

```{r}
depth_diff <- diff(depth_meas)
sum_pos_diff <- sum(depth_diff > 0)
  
```

Task 2: Consider sums of a three-measurement sliding window. How many sums are larger than the previous sum?

Let's calculate all three-measurement sums first:

```{r}
length_sums <- length(depth_meas)-2
three_sums <- depth_meas[1:length_sums] + depth_meas[2:(length_sums+1)] + depth_meas[3:(length_sums+2)]

```

Like before, calculate difference between subsequent rows and sum the positive ones:

```{r}
three_sums_diff <- diff(three_sums)
sum3_pos_diff <- sum(three_sums_diff > 0)
  
```

From twitter: sliding windows of length 3 will have the middle values in common, so you can just compare the values that enter/exit the window.
(And diff has a lag input!)

```{r}
sum3_pos_diff <- sum(diff(depth_meas,3)>0)

```

