---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
library(dplyr)
```

First, let's read in the input file:

```{r}
course <- read_table2("data/input2.txt", 
    col_names = c("direction","steps"), col_types = cols(steps = col_integer()))
```

Task 1: Calculate the horizontal position and depth you would have after following the planned course. What do you get if you multiply your final horizontal position by your final depth?

```{r}

course <- course %>%
  mutate(depth_step = case_when(direction == "down" ~ steps,
                              direction == "up" ~ -steps, 
                              direction == "forward" ~ as.integer(0)),
         horiz_step = ifelse(direction == "forward", steps, as.integer(0)))
         
destination <- course %>%
  summarise(hoz_pos = sum(horiz_step),
            final_dep = sum(depth_step),
            mult = hoz_pos * final_dep)

```

Task 2: In addition to horizontal position and depth, you'll also need to track a third value, aim, which also starts at 0. The commands also mean something entirely different than you first thought:

* down X increases your aim by X units.
* up X decreases your aim by X units.
* forward X does two things:
  * It increases your horizontal position by X units.
  * It increases your depth by your aim multiplied by X.

```{r}
course <- course %>%
  rename(aim = depth_step) %>%
  mutate(cum_aim = cumsum(aim),
         depth_step = cum_aim * horiz_step)

destination2 <- course %>%
  summarise(hoz_pos = sum(horiz_step),
            final_dep = sum(depth_step),
            mult = hoz_pos * final_dep)

```

